Alpine snow algae blooms vary with elevation

### 2020-01-11 manuscript/casey1.Rmd
responded to KY edits 1-10

---
title: "Alpine snow algae blooms vary with elevation"
author:
  - name: Casey B. Engstrom
    institute: SFU
  - name: Kurt M. Yakimovich
    institute: SFU
  - name: Lynne M. Quarmby
    institute: SFU
    email: quarmby@sfu.ca
institute:
  - SFU: Dept. Molecular Biology and Biochemistry, Simon Fraser University, 8888 University Drive, Burnaby, BC, V5A 1S6, Canada
output: 
  bookdown::word_document2:
    fig_caption: yes
    pandoc_args:
      - '--lua-filter=templates/scholarly-metadata.lua'
      - '--lua-filter=templates/author-info-blocks.lua'
    reference_docx: templates/word-template.docx
  # pdf_document:
  #   fig_caption: yes
  #   number_sections: yes
csl: the-isme-journal.csl
bibliography: refs.bib

@@

<!-- metabarcoding -->
In contrast to the high diversity of snow algae species observed in culture and Sanger sequencing, metabarcoding suggests that snow algae blooms are highly similar across continents, and are dominated by few taxa. 33 red snow communities across the Arctic were remarkably similar to each other: all were dominated by two operational taxonomic units (OTUs) of "uncultured Chlamydomonadaceae 1 and 2" with low relative abundance of *Chloromonas polyptera* and *Raphidonema nivale* [@lutz_biogeography_2016]. Similarly, another study using ITS2 metabarcoding found 24 red snow sites across the Arctic and Antarctic contained similar assemblages, dominated by "uncultured Chlamydomonadaceae A and B" with secondary abundance of *Raphidonema* and *Chloromonadinia* [@segawa_bipolar_2018]. Other 18S metabarcoding studies of snow algae were limited to genus or family level taxonomy [@hamilton_primary_2017]---because 18S is highly conserved, this marker generally cannot distinguish between closely related taxa.

<!-- our goal, what we did, what we found -->
Virtually nothing is known about the regional variation in species composition of snow algae blooms. Our goal was to identify the algae in blooms on different mountains and elevations in the Coast Range of British Columbia, Canada. We used 18S and *rbcL* metabarcoding to assess species composition---the latter targeting a hypervariable region of *rbcL* (coding for large subunit of rubisco), thus targeting only photosynthetic species with high taxonomic resolution. Additionally, we used light microscopy to describe the relative abundance of different morphologies in each field sample. By using these three cross-referenced metrics of relative abundance we were able to describe the algal species assemblage with high accuracy. Using *rbcL* as a marker revealed much greater diversity than would have been found using 18S alone. We did not find any differences in bloom composition between mountains, but there were distinct algae species assemblages at high versus low elevations.
<!-- stronger hook, predictions??? cliff hangers -->

@@

We collected snow algae from sites throughout the Coast Range near Vancouver, British Columbia, Canada over the summer of 2018 (Supplementary Figure S1). To capture the extent of snow algae diversity we sampled from different elevations, dates, mountains, and micro-habitats within the snow. In total, we collected 309 colored snow samples from a range of 880 - 2150 m elevation, including samples from 13 different mountains collected on 33 different dates (Supplementary Table S1). Our samples trended upwards in elevation over the course of the summer as the snow melted earliest at the lowest elevations (Supplementary Figure S2). We first detected snow algae above 1500 m on June 20. We scooped samples from visibly colored snow into sterile 50 mL centrifuge tubes. To prevent melting en route to the lab we stored tubes in a bag of snow. We melted each sample at room temperature on the bench and removed a 1 mL aliquot for light microscopy. Immediately after, samples were stored at -20 &deg;C for up to eight months until DNA extraction.

@@

We selected 33 out of our 310 samples for *rbcL* and 18S Illumina amplicon sequencing. We chose these samples to represent the variation in date, elevation, geographic location, snow color, micro-habitat, and cell morphology. We freeze-dried these samples for up to 48 hours until until samples appeared completely desiccated. We mini-pestled between 5 to 20 mg of sample at room temperature to physically rupture cell walls before incubation in CTAB lysis buffer [@noauthor_ctab_2009]. We extracted DNA using chloroform:isoamyl alcohol [@cubero_dna_1999], and spin columns (Qiagen, Hilden) (detailed protocol in Supplementary File S1). As a negative control we processed a sterile distilled water sample alongside each batch.

<!-- arent we following some extraction protocol here? cant we cite, and then have a supplementary protocol with our variation??? -->

<!-- To extract DNA from the crushed cells we added 800 mL  , 1% B-mercaptoethanol, 5 $\mu$L each of proteinase K and RNAase A, and incubated these at 65 &deg;C for 30 minutes. We spun samples at 10,000 g for 3 minutes to pellet cell debris, then added 700 $\mu$L of supernatant to an equal volume of 24:1 chloroform:isoamyl alcohol. We inverted samples to mix and centrifuged again at 12,000 g for 10 minutes. We precipitated DNA by transferring the top layer to 700 $\mu$L ice-cold ethanol, which we gently mixed, and spun over Qiagen DNA columns for 30 s at 15,000 g. We washed columns twice with 70% ethanol, and finally dissolved our DNA by spinning with 50 $\mu$L of sterile TE buffer.  -->

We designed custom primers to target a hypervariable region of snow algae *rbcL* (coding for the large subunit of rubisco). *rbcL* is more differentiated between snow algae than 18S (Supplementary Figure S3), and contains similar number of reference sequences on GenBank. Our *rbcL* primers target a 400 bp section of snow algae *rbcL* based off the consensus of 20 GenBank snow algae sequences (Supplementary Table S2).  We designed our primers with the Eurofins primer design tool (https://www.eurofinsgenomics.eu/en/ecom/tools/pcr-primer-design/). *Sanguina* sequences were not included because they were not available at the time. To target 18S, we used the universal primers Euk1181 and Euk1624 targeting the V7-V8 hypervariable regions [@wang_optimal_2014]. Primer sequences are available in Supplementary Table S3.

We constructed our 18S and *rbcL* amplicon libraries using a standard two-step PCR approach [@meyer_illumina_2010]. For detailed PCR conditions, see Supplementary File S2. After each PCR, we purified using Agencourt AMPure XP kit (Beckman Coulter, Brea, CA). We standardized sample DNA concentration using Qubit fluorometric quantification (Thermo Fisher, Waltham, MA). The pooled library was loaded and run on a MiSeq V3 kit (Illumina, San Diego, CA).

@@

We demultiplexed reads using CUTADAPT [@martin_cutadapt_2011], and filtered and trimmed, removed errors, dereplicated, merged paired-end reads, and removed chimeras following the default pipeline of DADA2 [@callahan_dada2_2016]. We assigned taxonomy to amplicon sequence variants (ASVs) using IDTaxa [@murali_idtaxa_2018], only assigning names to ASVs that were assigned with high confidence (threshold set to 50). We assigned taxonomy using custom snow algae reference databases for both 18S and *rbcL*, in addition to re-assigning 18S with SILVA to classify non-snow algae [@quast_silva_2013]. Snow algae reference databases and the scripts to generate them are available on our github repository (see below).

@@

To generate OTUs we clustered ASVs using t-SNE [@maaten_visualizing_2008]. We compared t-SNE results with the perplexity parameter ranging from 1 to 100, and found comparable results from 10 to 50. We categorized OTUs by applying DBSCAN [@hahsler_dbscan_2019] to our t-SNE. We compared *rbcL* sample similarity using UniFrac with default settings [@lozupone_unifrac_2005]. To confirm that our clustering was not an artifact of t-SNE we created maximum likelihood phylogenetic trees of ASVs using IQTree [@nguyen_iq-tree:_2015] (Supplementary Figures S4-S5).

All raw sequence data are available under European Nucleotide Archive accession PRJEB34539. All scripts used in this study are available at https://github.com/cengstro/bc_snow_algae_amplicon.

# Results {-}

@@

We observed morphologically distinct snow algae communities at different elevations (Figure 1). Red snow was prevalent in areas of high solar exposure above treeline, predominantly containing cells similar in appearance to *Sanguina nivaloides* [@prochazkova_sanguina_2019]. We occasionally observed green snow patches in alpine meadows receiving full sun, but more frequently observed green snow in well-shaded, forested sites. We found cells resembling *Chlainomonas rubra* [@novis_two_2008] at all elevations. Forested sites at low elevations often contained green or orange patches of snow, predominated by green or orange ovate cells with short spines resembling *Chloromonas cf. brevispina* [@matsuzaki_revisiting_2015].

Both 18S and *rbcL* amplicon libraries were dominated by ASVs that were assigned to Chlorophyta. Our 18S library contained 67 ASVs that were assigned to Chlorophyta and 7 ASVs to Ochrophyta, while our *rbcL* library contained 603 ASVs assigned to Chlorophyta and 41 to Trebouxiophyceae. Negative controls did not contain Qubit-detectable amounts of DNA. Our *rbcL* library contained higher ASV and OTU diversity than 18S: t-SNE clustering generated 9 *rbcL* OTUs, compared with just 4 18S OTUs (Supplementary Figure S6). The most frequent *rbcL* taxonomic assignments at genus level were *Chloromonas*, *Chlainomonas*, and *Sanguina*. Although most ASVs were not assigned to genus level, most ASVs clustered near one of these genera in NMDS ordination (Figure 2). Several *rbcL* OTUs were dominated by ASVs that were not assigned to genus level, including three OTUs closely related to *Chloromonas*, and one Chlamydomonadaceae OTU that was only assigned to the family level.

@@

Overall, our results show bloom composition can be highly variable on the same mountain, even in nearby patches. This is in contrast to previous papers that found snow algae assemblages were homogeneous across geographically distant sites [@lutz_biogeography_2016;@segawa_bipolar_2018]. This could be due to different sampling schemes: the aforementioned studies exclusively sampled high-latitude red snow, while our study included mid-latitude red, orange, and green snow from above and below treeline. Another recent study using metabarcoding found different snow algae communities in red and green snow algae blooms in Antarctica, including yellow blooms dominated by Chrysophyceae [@soto_microbial_2020]. Much of the diversity in our samples within *Chlainomonas* and *Chloromonas* was only detected with *rbcL* and not 18S, so the diversity within these genera may have been overlooked in previous metabarcoding studies.

Our *rbcL* data suggest *Chlainomonas* may be more common than previously thought. We could not differentiate between *Chlainomonas* and *Chloromonas* in our 18S data set, so this genus may have been overlooked in previous studies. *Chlainomonas* was previously thought to be restricted to waterlogged snow overlying mountain lakes [@novis_two_2008;@prochazkova_ecophysiological_2018], but we did not find any evidence that *Chlainomonas* was restricted in its habitat distribution. To the contrary, only one *Chlainomonas* dominant sample was located in waterlogged snow near the edge of a pool (sample S9); the other *Chlainomonas* dominant sites were not notably wetter than the surrounding snow, nor overlying frozen bodies of water. Interestingly, we found *Chlainomonas* was in higher relative abundance in our *rbcL* dataset than in cell counts (Figure 3)---possibly due to multiple chloroplasts per cell, which have been documented in this genus [@prochazkova_ecophysiological_2018].

Our findings highlight the remaining unexplored diversity in the snow algae microbiome. Many *rbcL* OTUs did not closely match any GenBank nucleotide archive sequences (Figure 2). "Chlamydomonadaceae E" highest BLAST percent identity was only 87%, and the top 10 BLAST matches contained 6 different algae genera---including two *Chloromonas* snow algae from Japan. This OTU did not correlate with any categories in our 18S or cell count data. One possible explanation is that it was lumped together with *Chloromonas* in 18S, and our crude quantification and categorization of cell counts did not detect this taxa. Other OTUs were also poorly matched to GenBank,  which could represent novel species of *Chloromonas* (Figure 2). The lack of clustering we observed in the remaining *Chloromonas* could be due to overlap in the levels of interspecific and intraspecific variation. *rbcL* diversity is likely higher within *Chloromonas* than in other algae genera: unlike *Sanguina* [@prochazkova_sanguina_2019], most species of *Chloromonas* lack a pyrenoid [@nozaki_differences_2002], which houses high concentrations of cross-linked rubisco. Likely related to its lack of pyrenoid, *Chloromonas* has many non-synonymous mutations in the region of *rbcL* the codes for binding rubisco together [@nozaki_differences_2002]. While *rbcL* is problematic in *Chloromonas* phylogeny [@nozaki_differences_2002], it nonetheless is highly differentiated and therefore an effective barcode for microalgae [@hall_assessment_2010].

@@

In conclusion, our study reveals substantial diversity within and between snow algae blooms. We found greater diversity using *rbcL* metabarcoding than 18S, which suggests snow algae diversity may have been overlooked in studies that used only 18S. Species were differentially distributed along an elevational gradient, with distinct blooms dominated by *Chloromonas*, *Sanguina*, or *Chlainomonas*. Future studies will examine other taxa present in snow algae blooms, and their interactions in these globally important microbiomes.

@@

# Author Contributions {-}

@@

# Additional information {-}

>>>>

---
title: "Alpine snow algae blooms vary with elevation"
author:
  - name: Casey B. Engstrom
    institute: SFU
  - name: Kurt M. Yakimovich
    institute: SFU
  - name: Lynne M. Quarmby
    institute: SFU
    email: quarmby@sfu.ca
institute:
  - SFU: Dept. Molecular Biology and Biochemistry, Simon Fraser University, 8888 University Drive, Burnaby, BC, V5A 1S6, Canada
output: 
  bookdown::word_document2:
    fig_caption: yes
    pandoc_args:
      - '--lua-filter=templates/scholarly-metadata.lua'
      - '--lua-filter=templates/author-info-blocks.lua'
    reference_docx: templates/word-template.docx
  # pdf_document:
  #   fig_caption: yes
  #   number_sections: yes
csl: templates/the-isme-journal.csl
bibliography: refs.bib

@@

<!-- metabarcoding -->
In contrast to the high diversity of snow algae species observed in culture, metabarcoding suggests that snow algae blooms are highly similar across continents, dominated by few taxa. One study of 33 red snow communities across the Arctic found all samples were dominated by two OTUs of "uncultured Chlamydomonadaceae 1 and 2" with low relative abundance of *Chloromonas polyptera* and *Raphidonema nivale* [@lutz_biogeography_2016]. Similarly, another study using ITS2 metabarcoding found 24 red snow sites across the Arctic and Antarctic contained similar assemblages, dominated by "uncultured Chlamydomonadaceae A and B" with secondary abundance of *Raphidonema* and *Chloromonadinia* [@segawa_bipolar_2018]. Other 18S metabarcoding studies of snow algae were limited to genus or family level taxonomy [@hamilton_primary_2017]---because 18S is highly conserved, this marker generally cannot distinguish between closely related taxa.

<!-- our goal, what we did, what we found -->
Virtually nothing is known about the regional variation in species composition of snow algae blooms, particularly in alpine regions. We set out to identify the algae in blooms on different mountains and elevations in the Coast Range of British Columbia, Canada. We used 18S and *rbcL* metabarcoding to assess species composition---the latter targeting a hypervariable region of *rbcL* (coding for large subunit of Rubisco), thus targeting only photosynthetic species with higher taxonomic resolution than 18S. We also used light microscopy to describe the relative abundance of different morphologies in each field sample. Using these three cross-referenced metrics we described the algal species assemblage in our study area with high accuracy. Using *rbcL* as a marker revealed much greater diversity than would have been found using 18S alone. We did not find any differences in bloom composition between mountains, but there were distinct algae species assemblages at high versus low elevations.

@@

<!-- the problem with this is we cherry picked our samples. We weren't doing transects or anything like that. A reviewer could say, you couldve picked your samples that supported your hypothesis about different bloom types. which is why I brought in the cell count data. -->

We collected snow algae from sites throughout the Coast Range near Vancouver, British Columbia, Canada over the summer of 2018 (Supplementary Figure S1). To capture the extent of snow algae diversity we sampled from different elevations, dates, mountains, and micro-habitats within the snow. In total, we collected 309 colored snow samples from a range of 880 - 2150 m elevation, including samples from 13 different mountains collected on 33 different dates (Supplementary Table S1). Our samples trended upwards in elevation over the course of the summer as the snow melted earliest at the lowest elevations. We first detected snow algae above 1500 m on June 20. We scooped samples from visibly colored snow into sterile 50 mL centrifuge tubes. To prevent melting en route to the lab we stored tubes in a bag of snow. We melted each sample at room temperature on the bench and removed a 1 mL aliquot for light microscopy. Immediately after, samples were stored at -20 &deg;C for up to eight months until DNA extraction.

@@

We selected 33 out of our 310 samples for *rbcL* and 18S Illumina amplicon sequencing. We chose these samples to represent the variation in date, elevation, geographic location, snow color, micro-habitat, and cell morphology. We freeze-dried these samples for up to 48 hours until until samples appeared completely desiccated. We mini-pestled between 5 to 20 mg of sample at room temperature to physically rupture cell walls before incubation in CTAB lysis buffer [@noauthor_ctab_2009]. We extracted DNA using chloroform:isoamyl alcohol [@cubero_dna_1999], and spin columns (Qiagen, Hilden) (details available in Supplementary Protocol S1). As a negative control we processed a sterile distilled water sample alongside each batch.

We designed custom primers to target a hypervariable region of snow algae *rbcL*. *rbcL* is more differentiated between snow algae than 18S (Supplementary Figure S2), and contains similar number of reference sequences on GenBank. Our *rbcL* primers target a 400 bp section of snow algae *rbcL* based off the consensus of 20 GenBank snow algae sequences (Supplementary Figure S2). We designed our primers with the Eurofins primer design tool (https://www.eurofinsgenomics.eu/en/ecom/tools/pcr-primer-design/). *Sanguina* sequences were not included because they were not available at the time. To target 18S, we used the universal primers Euk1181 and Euk1624 targeting the V7-V8 hypervariable regions [@wang_optimal_2014]. Primer sequences are available in Supplementary Table S2.

We constructed our 18S and *rbcL* amplicon libraries using a standard two-step PCR approach [@meyer_illumina_2010]. For detailed PCR conditions, see Supplementary Protocol S2. After each PCR, we purified using Agencourt AMPure XP kit (Beckman Coulter, Brea, CA). We standardized sample DNA concentration using Qubit fluorometric quantification (Thermo Fisher, Waltham, MA). The pooled library was loaded and run on a MiSeq V3 kit (Illumina, San Diego, CA).

@@

We demultiplexed reads using CUTADAPT [@martin_cutadapt_2011], and filtered and trimmed, removed errors, dereplicated, merged paired-end reads, and removed chimeras following the default pipeline of DADA2 [@callahan_dada2_2016]. We assigned taxonomy to amplicon sequence variants (ASVs) using IDTaxa [@murali_idtaxa_2018], only assigning names to ASVs that were assigned with high confidence (threshold set to 50). We assigned taxonomy using custom snow algae reference databases for both 18S and *rbcL*, in addition to re-assigning 18S with SILVA to classify non-snow algae [@quast_silva_2013]. Snow algae reference databases and the scripts to generate them are available on our github repository.

@@

To generate OTUs we clustered ASVs using t-SNE [@maaten_visualizing_2008]. We compared t-SNE results with the perplexity parameter ranging from 1 to 100, and found comparable results from 10 to 50. We categorized OTUs by applying DBSCAN [@hahsler_dbscan_2019] to our t-SNE. We compared *rbcL* sample similarity using UniFrac with default settings [@lozupone_unifrac_2005]. To confirm that our clustering was not an artifact of t-SNE we created maximum likelihood phylogenetic trees of ASVs using IQTree [@nguyen_iq-tree:_2015] (Supplementary Figure S3).

# Results {-}
<!-- should we report the diversity data quickly? or tangential to our question... -->

@@

We observed morphologically distinct snow algae communities at different elevations (Figure 1; Supplementary Figure S4). Red snow was prevalent in areas of high solar exposure above treeline, predominantly containing cells similar in appearance to *Sanguina nivaloides* [@prochazkova_sanguina_2019]. We occasionally observed green snow patches in alpine meadows receiving full sun, but more frequently observed green snow in well-shaded, forested sites. We found cells resembling *Chlainomonas rubra* [@novis_two_2008] at all elevations. Forested sites at low elevations often contained green or orange patches of snow, predominated by green or orange ovate cells with short spines resembling *Chloromonas cf. brevispina* [@matsuzaki_revisiting_2015].

Both 18S and *rbcL* amplicon libraries were dominated by ASVs that were assigned to Chlorophyta. Our 18S library contained 67 ASVs that were assigned to Chlorophyta and 7 ASVs to Ochrophyta, while our *rbcL* library contained 603 ASVs assigned to Chlorophyta and 41 to Trebouxiophyceae. Negative controls did not contain Qubit-detectable amounts of DNA. Our *rbcL* library contained higher ASV and OTU diversity than 18S, t-SNE clustering generated 9 *rbcL* OTUs, while we detected just 3 visually distinct clusters among 18S ASVs (Supplementary Figure S5). For *rbcL*, the most frequent taxonomic assignments at genus level were *Chloromonas*, *Chlainomonas*, and *Sanguina*. Although most ASVs were not assigned to genus level, most ASVs clustered near one of these genera in NMDS ordination (Figure 2). Several OTUs were dominated by ASVs that were not assigned to genus level, including three OTUs closely related to *Chloromonas*, and one Chlamydomonadaceae OTU that was only assigned to the family level.

@@

Our results show bloom composition can be highly variable on the same mountain, even in nearby patches. In contrast, previous studies found snow algae assemblages were homogeneous across geographically distant sites [@lutz_biogeography_2016;@segawa_bipolar_2018]. The aforementioned studies exclusively sampled high-latitude red snow, while our study included mid-latitude red, orange, and green snow from above and below treeline. Another recent study using 18S metabarcoding found different snow algae communities in red and green snow algae blooms in Antarctica, including yellow blooms dominated by Chrysophyceae [@soto_microbial_2020]. Much of the diversity in our samples within *Chlainomonas* and *Chloromonas* was only detected with *rbcL* and not 18S, so the diversity within these genera may have been overlooked in previous metabarcoding studies.

Our *rbcL* data suggest *Chlainomonas* may be more widely distributed than previously thought. We could not differentiate between *Chlainomonas* and *Chloromonas* in our 18S data set, so this genus may have been overlooked in previous studies. *Chlainomonas* was previously thought to be restricted to waterlogged snow overlying mountain lakes [@novis_two_2008;@prochazkova_ecophysiological_2018], but we did not find any evidence that *Chlainomonas* was restricted in its habitat distribution. To the contrary, only one *Chlainomonas* dominant sample was located in waterlogged snow near the edge of a pool (sample S9); the other *Chlainomonas* dominant sites were not notably wetter than the surrounding snow, nor overlying frozen bodies of water. Interestingly, we found *Chlainomonas* was in higher relative abundance in our *rbcL* dataset than in cell counts (Figure 3)---possibly due to multiple chloroplasts per cell, which have been documented in this genus [@prochazkova_ecophysiological_2018].

Our findings highlight the remaining unexplored diversity in the snow algae microbiome. Many *rbcL* OTUs did not closely match any GenBank nucleotide archive sequences (Figure 2). "Chlamydomonadaceae E" highest BLAST percent identity was only 87%, and the top 10 BLAST matches contained 6 different algae genera---including two *Chloromonas* snow algae from Japan. This OTU did not correlate with any categories in our 18S or cell count data. One possible explanation is that it was lumped together with *Chloromonas* in 18S, and our crude quantification and categorization of cell counts did not detect this taxa. Other OTUs were also poorly matched to GenBank,  which could represent novel species of *Chloromonas* (Figure 2). The lack of clustering we observed in the remaining *Chloromonas* could be due to overlap in the levels of interspecific and intraspecific variation. *rbcL* diversity is likely higher within *Chloromonas* than in other algae genera: unlike *Sanguina* [@prochazkova_sanguina_2019], most species of *Chloromonas* lack a pyrenoid [@nozaki_differences_2002], which houses high concentrations of cross-linked Rubisco. Likely related to its lack of pyrenoid, *Chloromonas* has many non-synonymous mutations in the region of *rbcL* the codes for binding Rubisco together [@nozaki_differences_2002]. While *rbcL* is problematic in *Chloromonas* phylogeny [@nozaki_differences_2002], it nonetheless is highly differentiated and therefore an effective barcode for microalgae [@hall_assessment_2010].

@@

In conclusion, our study reveals substantial diversity within and between snow algae blooms. We found greater diversity using *rbcL* metabarcoding than 18S, which suggests snow algae diversity may be overlooked in studies using only 18S. Species were differentially distributed along an elevational gradient, with distinct blooms dominated by *Chloromonas*, *Sanguina*, or *Chlainomonas*. Future studies will examine other taxa present in snow algae blooms, and their interactions in these globally important microbiomes.

<!-- This needs to be beefed up. Some sort of statement about how this work impacts the field, more specifics about future directions. How this work could inform albedo work that you mention in your intro.  -->

# Data availability {-}

All raw sequence data are available under European Nucleotide Archive accession PRJEB34539. All scripts used in this study are available at https://github.com/cengstro/bc_snow_algae_amplicon. Supplementary materials are available at https://www.nature.com/ismej/.

@@

# Author contributions {-}

@@

# Conflict of interest {-}
